Effectful Semantics for Substructural SSA

### 2023-11-16 pldi24/paper.tex
Microtap

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\isrw{\mc{H}}{I}{H}\) & \(\mc{H}\) is a rewrite from \(I\) to \(H\),
        i.e. fills every hole in \(H\) modulo \(I\) \\
        \(\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\), \(\hasty{\Gamma}{p}{[\mc{H}]\mhole{a}}{A}{q}\) \\
        \(\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\), \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
        \(\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\),
        \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Typing judgements for \isotopessa terms, blocks, and CFGs with holes}
  \Description{Typing judgements for isotope-SSA terms, blocks, and CFGs with
  holes}
  \label{fig:ssa-holes-judgements}
\end{figure}
The typing rules for terms, blocks, and CFGs with holes are exactly the same as
the corresponding rules without holes, modulo the additional rules
\rle{hole-tm}, \rle{hole-blk}, and \rle{hole-cfg} to type the holes themselves;
they are given in full in Figures \ref{fig:ssa-term-holes-typing} and
\ref{fig:ssa-block-holes-typing}.

@@

This formalism allows us to state the \textit{syntactic rewriting} theorem,
which says, in essence, that filling holes with valid terms yields valid terms.
\begin{theorem}[Rewriting]
  Given \(\isrw{\mc{H}}{H}{I}\), we have
  \begin{itemize}
    \item If \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), then \(\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}\)
    \item If \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), then \(\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}\)
  \end{itemize}
  Furthermore, we have that 
  \begin{itemize}
    \item \(\mhasty{\cdot}{\Gamma}{p}{a}{A}{q} \iff \hasty{\Gamma}{p}{a}{A}{q}\)
    \item \(\mhaslb{\cdot}{\Gamma}{t}{\ms{L}} \iff \haslb{\Gamma}{t}{\ms{L}}\)
  \end{itemize}
\end{theorem}

>>>>

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\isrw{\mc{H}}{I}{H}\) & \(\mc{H}\) is a rewrite from \(I\) to \(H\),
        i.e. fills every hole in \(H\) modulo \(I\) \\
        \(\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\), \(\hasty{\Gamma}{p}{[\mc{H}]\mhole{a}}{A}{q}\) \\
        \(\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\), \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
        \(\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}\) & Given
        \(\isrw{\mc{H}}{\cdot}{H}\),
        \(\haslb{\Gamma}{[\mc{H}]\mhole{t}}{\ms{L}}\) \\
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Typing judgements for \isotopessa terms, blocks, and CFGs with holes}
  \Description{Typing judgements for isotope-SSA terms, blocks, and CFGs with
  holes}
  \label{fig:ssa-holes-judgements}
\end{figure}

@@

The typing rules for terms, blocks, and CFGs with holes are exactly the same as
the corresponding rules without holes, modulo the additional rules
\rle{hole-tm}, \rle{hole-blk}, and \rle{hole-cfg} to type the holes themselves;
they are given in full in Figures \ref{fig:ssa-term-holes-typing} and
\ref{fig:ssa-block-holes-typing}. This formalism allows us to state the
\textit{syntactic rewriting} theorem, which says, in essence, that filling holes
with valid terms yields valid terms.
\begin{theorem}[Rewriting]
  Given \(\isrw{\mc{H}}{H}{I}\), we have
  \begin{itemize}
    \item If \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), then \(\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}\)
    \item If \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), then \(\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}\)
  \end{itemize}
  Furthermore, we have that 
  \begin{itemize}
    \item \(\mhasty{\cdot}{\Gamma}{p}{a}{A}{q} \iff \hasty{\Gamma}{p}{a}{A}{q}\)
    \item \(\mhaslb{\cdot}{\Gamma}{t}{\ms{L}} \iff \haslb{\Gamma}{t}{\ms{L}}\)
  \end{itemize}
\end{theorem}
